%        File: dcg.tex
%     Created: Fri Oct 25 11:00 AM 2013 C
% Last Change: Fri Oct 25 11:00 AM 2013 C
%
\documentclass[a4paper]{article}

\usepackage[utf8]{inputenc}

\usepackage{amsmath}
\usepackage{amsfonts}

\begin{document}

\section{Discounted cumulative gain}

Discounted cumulative gain is a measure of the quality of a search engine, in information retrieval.

\subsection{Cumulative gain}
Let $\left\{ X_1, \ldots, X_N \right\}$ be the first $N$ results of a given query and let $rel(\cdot)$ 
be the associated relevance function. The cumulative gain at rank $p \le N$ is defined by:
\[
  \mathrm{CG}_p = \sum_{i = 1}^{p} rel(X_i).
\]
It is simply the cumulated relevance of the first $p$ results.

\subsection{Discounted cumulative gain}
Cumulative gain does not take into account the fact that a relevant result should be penalized if it 
appears at a large rank. The discounted cumulative gain is based on the cumulative gain definition 
but the relevance is penalized logarithmically proportional to the rank of the result. There are two 
definitions of the discounted cumulative gain. Even if they are quite similar, they are not equivalent. 
The first definition is given by:
\[
  \mathrm{DGC}_p = rel(X_1) + \sum_{i = 2}^{p} \frac{rel(X_i)}{\log_2(i)}.
\]
The second definition is given by:
\[
  \mathrm{DGC}_p = \sum_{i = 1}^{p} \frac{2^{rel(X_i)} - 1}{\log_2(i + 1)}.
\]
%TODO: check theoretical explanation about log_2 on denominator

\subsection{Normalized discounted cumulative gain}
DCG can be useful to characterize a search engine's performance in response to one query. However, 
it does not allow for consistent comparison for different queries: the length of the results list 
can vary and this is not taken into account by DCG alone. Let $\mathrm{IDCG}_p$ be defined for 
$p \le N$ as follow:
\[
  \mathrm{IDCG}_p = \max_{\sigma \in \mathfrak{S}} DCG_p\left( \left( X_{\sigma(1)}, \ldots, X_{\sigma(p)} \right) \right).
\]
This represents the best DCG possible considering $p$ results from the full list. We then define 
the normalized discounted cumulative gain by:
\[
  nDCG_p = \frac{DCG_p}{IDCG_p}.
\]
The nDCG represents the quality of the $p$ firsts results, compared to the $p$ best results from the
full set.

Normalized discounted cumulative gain can be very useful when addressing a ranking problem for a recommender.
The problem is then to sort recommandations and to show the top-$N$ ones, where $N$ is very small compared 
to the total number of possible recommendations. Hence, it is important that top recommendations are more 
emphasized than other ones.

\section{Mean reciprocal rank}
The mean reciprocal rank is a statistic used for evaluating a process producing a 
list of possible answers to a given query. The smallest the rank of the correct 
answer, the higher the MRR.
Let $\mathcal{Q}$ be a set of query and let $\left( r_q \right)_{q \in \mathcal{Q}}$ 
be the corresponding ranks of the correct answers. The MRR associated to $\mathcal{Q}$ 
is defined by:
\[
  \mathrm{MRR}(\mathcal{Q}) = \frac{1}{|\mathcal{Q}|} \sum_{q \in \mathcal{Q}} \frac{1}{r_q}.
\]

\end{document}
